---
title: "General Social Survey"
author: "Marshette Williams"
date: "April 27,2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"
mydata5 <- filter(mydata4, CAPPUN=="FAVOR" | CAPPUN=="OPPOSE")
```




# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. First explain what you make out of comparing how people voted in 2012 with 2016.


How are the "voted in 2016" and the "voted in 2012" related? Among the people who voted in 2012 shows consistence with the people who voted in 2016
$H_0:$ voting in 2016 and 2012  are unrelated


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
 Use catergorical varibles, and a chi-squared test to test hypothesis that voting in 2016 is / is not independenbt of voting in 2012
##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

we use a barchart to show the voting trends between 2012 and 2016

#### Graphical Descriptive Results
```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent") 

```


<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

 numerical summary is a cross tabulation:
```{r}
xtabs(~VOTE16+VOTE12, data=mydata4)
```


<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

whether you voted in 2012 is related to voting in 2016

```{r}
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
fisher.test(votetab)
```

They hypothesis voting in 2012 is related to voting 2016 which we come to the conclusin is the p value is 0 and it Reject the null hypothesis

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
 
The hypothesis was that voting in 2012 is 2016 related. We used a chi squared test of independence the p value is equal to 0. Therefore the null hypothesis is related. We reject the null hypothesis because the p value is less then 0.5 we have evidence that voting in 2012 is related to voting in 2016 and if you do not vote in 2012 yo likely did not vote in 2016.

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->


Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

$H_0:$ Do you believe bigbang theory and afterlife. Are the two related?

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Use catergorical varibles, and a chi-squared test to test hypothesis that voting in BIGBANG is / is not independenbt of POSTLIFE

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

we use a barchart to show the relation between the big bang and postlife


#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata3, type = "percent",xlab="BELIEVE IN POSTLIFE",main='BELIEVE IN BIGBANG')

```



#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
 numerical summary is a cross tabulation: bIGBANG AND POSTLIFE
```{r}

xtabs(~BIGBANG+POSTLIFE, data=mydata3)
```
 

 
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
whether you believe in the bigbang and postlife

```{r}
chisqtestGC(~BIGBANG+POSTLIFE, data=mydata3)
BANGTABLE=xtabs(~BIGBANG+POSTLIFE, data=mydata3)
fisher.test(BANGTABLE)
```
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
The hypothesis was the the belief in the big bang thoery is related to the belief in the afterlife. We used a chi squared test of independence, because the p value is less then 0.5 meaning the null hypothesis is related the null since we have evidence that the belief i the postive is related to the belief in the big bang thoery and if you do not believe in the big bang theory.



# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This time compare the opinion on death penalty (CAPPUN) with opinion on courts (COURTS).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Use catergorical varibles, and a chi-squared test to test hypothesis of CAPPUN is / is not independent of COURTS

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results
we use a barchart to show the opinion of Cappun and how they feel regarding the courts


#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN+COURTS, data=mydata5, type = "percent",xlab="VIEW ON CAPPUN",main='VIEW ON COURTS')
```



#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
xtabs(~CAPPUN+COURTS, data=mydata4)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~CAPPUN+COURTS, data=mydata4)
viewstab=xtabs(~CAPPUN+COURTS, data=mydata4)
fisher.test(viewstab)
```




# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The hypothesis was the voting in 2012 is related. We used a chi-squared test if independence the p value is equal 0. Therefore, the null hypothesis is related. We reject the null hypothesis because the p valve is less then .5. Meaning the null hypothesis is related. We rejected the null since we have evidence that the belief in the death penalty is related to the belief in the court system and if you do not believe in the death penalty it is likely you did not believe in the court system.

